### 2015-05-24 implementation.tex
Bunch of steps.

\subsection{OpenVPN}
% By default, point-to-point, but since the version 2.0 can be configured in server mode, hence being able to manage hundreds of clients.

@@

>>>>

\subsection{OpenVPN}
% By default, point-to-point, but since the version 2.0 can be configured in server mode, hence being able to manage hundreds of clients.
% TODO talk about the cipher none patch for version < ...

OpenVPN is widely used, and is can even integraed in some routers using packages such as DD-WRT and OpenWRT.
OpenVPN is also used by the Dutch government to secure all its communications~\cite{openvpn-nl}.

OpenVPN offers two different TUN/TAP virtual interfaces:
\begin{description}
	\item[TUN] Layer 3 IP tunnel
	\item[TAP] Layer 2 ethernet tunnel.
\end{description}

Uses regular TCP/UDP network protocols, which can be an advantage over IPSec if the ISP decides to block the latter.

Note that if one wants to use the \texttt{none} cipher with a version of OpenVPN ealier than 2.6.7, he would have to update it using a community patch~\cite{openvpn-patch-none}.

@@

Offers an high-level interface called EVP to be used by other applications.

### 2015-05-24 technical-background.tex
ipsec transport/tunnel overhead figure.
implementation introduction
patch openssh, implem openssh and drivers.
Bunch of steps.

\chapter{Technical background}

@@

% Present the TCP/IP layering

@@

There exist several major implementations of VPN: SSL, IPSec and PPTP.
The later was developped by a vendor consortium and proposed in the RFC 2637 and will not be discussed further.

\subsection{SSL/TLS}
% Introduce SSL/TLS, talk about the protocol, the key exchange and stuff, but leaver OpenVPN for the 'implementation' chapter.
% Question the security? Apparently, it does mac-then-encrypt, which is insecure regarding certain types of attacks \cite{cryptoeprint:2001:045}.
% Plus, SSLv3 is to be deprecated, according to a queued RFC: http://www.rfc-editor.org/internet-drafts/draft-ietf-tls-sslv3-diediedie-03.txt

@@

>>>>

\chapter{Technical background}\label{chap:theory}

@@

% Present the TCP/IP layering
The RFC 1122~\cite{rfc1122} defines the TCP/IP and OSI stack as in the table~\ref{tab:tcp-ip-stack}.
The main difference between the two is the application layer of the TCP/IP stack which corresponds to the three upper layers of the OSI model.
Some references, such as~\citet{tanenbaum2011}, conceptually split the TCP/IP link layer into an additional physical layer.

\begin{table}[ht]
\center
\begin{tabularx}{\textwidth}{|l|l|l|X|} \hline
\multicolumn{2}{|c|}{TCP/IP layering} & \multicolumn{2}{c|}{OSI model} \\ \hline
Layer & Protocols & Layer & Protocols \\ \hline
\multirow{3}{*}{Application} & \multirow{3}{*}{FTP, SSH} & Application & FTP \\ \cline{3-4}
 & & Presentation & ASCII, JPEG \\ \cline{3-4}
 & & Session & RPC, PAP \\ \hline
Transport & TCP, UDP & Transport & TCP, UDP \\ \hline
Internet & IP, ICMP, IPSec & Network & IP, ICMP, IPSec \\ \hline
\multirow{2}{*}{Data Link} & \multirow{2}{*}{PPP, MAC, Ethernet, L2TP} & Data link & PPP, MAC, Ethernet, L2TP \\ \cline{3-4}
 & & Physical & USB, DSL, IEEE 802.11 \\ \hline
\end{tabularx}
\caption{TCP/IP and OSI model comparison}{They are globaly the same, except for the application layer of the TCP/IP stack which merge togheter the three upper layers of the OSI model. Between parenthsis are examples of protocols resting on each layer.}
\label{tab:tcp-ip-stack}
\end{table}

@@

There exist several major implementations of VPN: SSL, IPSec, L2TP and PPTP.
The later was developped by a vendor consortium and proposed in the RFC 2637 and will not be discussed further.

\subsection{SSL/TLS}
% Introduce SSL/TLS, talk about the protocol, the key exchange and stuff, but leaver OpenVPN for the 'implementation' chapter.
% Question the security? Apparently, it does mac-then-encrypt, which is insecure regarding certain types of attacks \cite{cryptoeprint:2001:045}.
% Plus, SSLv3 is to be deprecated, according to a queued RFC: http://www.rfc-editor.org/internet-drafts/draft-ietf-tls-sslv3-diediedie-03.txt, and is not supported anymore by many servers since poodle.

@@

%TODO Figure IPSec frame structure.

\begin{figure}[ht]
\center
\subfloat[\label{fig:ipsec-transport}Transport]{%
	\large
	\resizebox{.465\linewidth}{!}{%
	\input{ipsec-transport}
	}
}
\subfloat[\label{fig:ipsec-tunnel}Tunnel]{%
	\large
	\resizebox{.535\linewidth}{!}{%
	\input{ipsec-tunnel}
	}
}
\caption{IPSec transport (a) and tunnel (b) overheads}{Tunnel mode adds a custom IP header and moves the AH/ESP header in front of the original IP header. Note: ``trl" stands for ``trailer", ``hdr" for ``header"}
\label{fig:ipsec-transport-tunnel}
\end{figure}
